Clinical prediction with localized modeling using similarity-based cohorts: A scoping review

### 2023-10-28 docs/Manuscript/Manuscript.Rmd
add table of specializations from general framework to studies

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
Sys.sleep(10)
tab_synthesis <- readr::read_rds(here::here("data/synthesis.rds"))
composite_studies <- readr::read_rds(here::here("data/composite.rds"))
tab_eval_comp <- readr::read_rds(here::here("data/eval-comp.rds"))
citation_id <- readr::read_rds(here::here("data/citations.rds"))
```

@@

We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018]. While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; clinical decision support and software tools). For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.

@@

```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
composite_studies |> 
  transmute(
    # Citation = str_c("@", Citation),
    Citation = sapply(
      str_split(Citation, "; "),
      function(s) str_c("@", s, collapse = "; ")
    ),
    Elements = str_replace_all(Elements, ",", ";"),
    Terminology
  ) |> 
  mutate(across(
    everything(),
    # ~ str_replace_all(., "<U+001B>", "")
    textclean::replace_non_ascii
  )) |>
  # mutate(across(c(Citation), ~ str_c(., "\\hspace{6em}"))) |> 
  knitr::kable(
    align = "lllllrr",
    caption = paste0(
      "\\label{tab:composite}",
      "Methodological elements of studies included in the synthesis. "
    )
  )
```

@@

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
Sys.sleep(10)
tab_synthesis <- readr::read_rds(here::here("data/synthesis.rds"))
composite_studies <- readr::read_rds(here::here("data/composite.rds"))
framework_specs <- readr::read_rds(here::here("data/framework.rds"))
tab_eval_comp <- readr::read_rds(here::here("data/eval-comp.rds"))
citation_id <- readr::read_rds(here::here("data/citations.rds"))
```

@@

We refer the interested reader to several previous reviews of CBR in medicine [@Gierl1998a; @Begum2011; @Choudhury2016], of measures of patient similarity [@Dai2020], and of uses of patient similarity in predictive models [@Welch2013; @Sharafoddini2017; @Parimbelli2018]. While the reviews of CBR focus on applications using health data, the patient similarity reviews encompass many additional types of data (various molecular -omics, genetic tests, medical images, laboratory tests, patient preferences, patient-reported outcomes, tracking devices, social media) and survey a much broader scope of models (exploratory analysis via dimension reduction and cluster analysis; risk evaluation and outcome prediction; clinical decision support and software tools). For the present review, we are interested in how similarity matching on patient-level health data is used to construct local cohorts for predictive or inferential modeling.

@@

```{r tab:composite, echo=FALSE, message=FALSE, warning=FALSE}
composite_studies |> 
  transmute(
    # Citation = str_c("@", Citation),
    Citation = sapply(
      str_split(Citation, "; "),
      function(s) str_c("@", s, collapse = "; ")
    ),
    Elements = str_replace_all(Elements, ",", ";"),
    Terminology
  ) |> 
  mutate(across(
    everything(),
    # ~ str_replace_all(., "<U+001B>", "")
    textclean::replace_non_ascii
  )) |>
  # mutate(across(c(Citation), ~ str_c(., "\\hspace{6em}"))) |> 
  knitr::kable(
    align = "lllllrr",
    caption = paste0(
      "\\label{tab:composite}",
      "Methodological elements of studies included in the synthesis."
    )
  )
```

@@

```{r tab:framework, echo=FALSE, message=FALSE, warning=FALSE}
framework_specs |> 
  mutate(
    Citation = sapply(
      str_split(Citation, "; "),
      function(s) str_c("@", s, collapse = "; ")
    )
  ) |> 
  mutate(Citation = str_c("[", Citation, "]")) |> 
  # prevent numeric list columns from dominating the horizontal space
  mutate(across(c(-Citation), ~ str_c(., "\\hspace{18em}"))) |> 
  knitr::kable(
    align = "llll",
    caption = paste0(
      "\\label{tab:framework}",
      "Specializations of general framework to included studies.",
      " Flags: R = recurse, S = supervised, T = tuned (optimized)."
    )
  )
```

### 2023-10-28 docs/Manuscript/Manuscript.tex
add table of specializations from general framework to studies

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{Gierl L, Bull M, Schmidt R. {CBR} in {Medicine}. In:
Lenz M, Burkhard H-D, Bartsch-Spörl B, Wess S, editors. Case-{Based
Reasoning Technology}, vol. 1400, {Berlin, Heidelberg}: {Springer Berlin
Heidelberg}; 1998, p. 273--97.
\url{https://doi.org/10.1007/3-540-69351-3_11}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Dai2020}{}}%
\CSLLeftMargin{{[}7{]} }%
\CSLRightInline{Dai L, Zhu H, Liu D.
\href{https://arxiv.org/abs/2012.01976}{Patient similarity: Methods and
applications}. arXiv:201201976 {[}Cs{]} 2020.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
\CSLLeftMargin{{[}39{]} }%
\CSLRightInline{Biecek P, Burzykowski T. Explanatory {Model Analysis}:
{Explore}, {Explain}, and {Examine Predictive Models}. {New York}:
{Chapman and Hall/CRC}; 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
\CSLLeftMargin{{[}40{]} }%
\CSLRightInline{Molnar C. Interpretable {Machine Learning}: {A Guide}
for {Making Black Box Models Explainable}. 2023.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
\CSLLeftMargin{{[}43{]} }%
\CSLRightInline{Yang L. Distance {Metric Learning}: {A Comprehensive
Survey}. PhD thesis. Michigan State University, 2006.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
\CSLLeftMargin{{[}45{]} }%
\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A. Distance {Metric
Learning} with {Application} to {Clustering} with {Side-Information}.
Advances in {Neural Information Processing Systems}, vol. 15, {MIT
Press}; 2002.}

>>>>

@@

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

@@

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2087}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2795}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1535}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3583}}@{}}
\caption{\label{tab:framework}Specializations of general framework to
included studies. Flags: R = recurse, S = supervised, T = tuned
(optimized).}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Relevance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Retrieval
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptation
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Citation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Relevance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Retrieval
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Adaptation
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
{[}11{]} & Levenshtein distance\hspace{18em} & Statistical\hspace{18em}
& Survival (S)\hspace{18em} \\
{[}19{]} & Weighted sum (S)\hspace{18em} & Combinatorial\hspace{18em} &
Proportion (S)\hspace{18em} \\
{[}20{]} & Kohonen mapping\hspace{18em} & Adaptive\hspace{18em} &
Representation (S)\hspace{18em} \\
{[}13{]} & Euclidean distance\hspace{18em} & Statistical
(T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
{[}22{]} & Entropic distance\hspace{18em} & Combinatorial\hspace{18em} &
Proportion (S)\hspace{18em} \\
{[}23{]} & Euclidean distance\hspace{18em} & Combinatorial\hspace{18em}
& Proportion (S)\hspace{18em} \\
{[}18,21,25,26{]} & Feature selection (R,S)\hspace{18em} & Combinatorial
(R)\hspace{18em} & Fuzzy rules (S,T)\hspace{18em} \\
{[}24{]} & User-defined\hspace{18em} & User-defined\hspace{18em} &
User-defined\hspace{18em} \\
{[}12{]} & Mahalanobis distance (S)\hspace{18em} & Combinatorial
(T)\hspace{18em} & Proportional hazards (S)\hspace{18em} \\
{[}27{]} & Weighted overlap\hspace{18em} & Combinatorial
(T)\hspace{18em} & Weighted sum (S)\hspace{18em} \\
{[}28{]} & Distance metric learning (S)\hspace{18em} &
Combinatorial\hspace{18em} & Association rules (S,T)\hspace{18em} \\
{[}15{]} & Locally supervised metric learning (S); Feature
selection\hspace{18em} & Combinatorial\hspace{18em} & Logistic model
(S)\hspace{18em} \\
{[}14{]} & Cosine similarity\hspace{18em} & Combinatorial
(T)\hspace{18em} & Logistic model (S); Decision tree (S)\hspace{18em} \\
{[}29{]} & Range overlap\hspace{18em} & Clustering (S)\hspace{18em} &
Artificial neural network (S)\hspace{18em} \\
{[}16{]} & Cosine similarity; Random forest proximity (S)\hspace{18em} &
Combinatorial (T)\hspace{18em} & Logistic model (S); Decision tree (S);
Random forest (S)\hspace{18em} \\
{[}31{]} & Unclear\hspace{18em} & Unclear\hspace{18em} &
Unclear\hspace{18em} \\
{[}30{]} & Gaussian process model\hspace{18em} & Complete\hspace{18em} &
Weighted support vector machine (S); Weighted Gaussian process
regression (S)\hspace{18em} \\
{[}17{]} & Weighted sum\hspace{18em} & Combinatorial (T)\hspace{18em} &
Logistic model (S); Random forest (S); Nearest neighbors
(S)\hspace{18em} \\
{[}32{]} & Weighted sum\hspace{18em} & Combinatorial\hspace{18em} &
Extreme learning machine (S)\hspace{18em} \\
{[}33{]} & Weighted sum\hspace{18em} & Statistical\hspace{18em} & Linear
model\hspace{18em} \\
{[}34,35{]} & Locally supervised metric learning (S)\hspace{18em} &
Causal inference matching\hspace{18em} & Statistical test
(S)\hspace{18em} \\
\end{longtable}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Gierl1998a}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{Gierl L, Bull M, Schmidt R. {CBR} in {Medicine}. In:
Lenz M, Burkhard H-D, Bartsch-Spörl B, Wess S, editors. Case-{Based
Reasoning Technology}, vol. 1400, {Berlin, Heidelberg}: {Springer Berlin
Heidelberg}; 1998, p. 273--97.
\url{https://doi.org/10.1007/3-540-69351-3_11}.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Dai2020}{}}%
\CSLLeftMargin{{[}7{]} }%
\CSLRightInline{Dai L, Zhu H, Liu D.
\href{http://arxiv.org/abs/2012.01976}{Patient similarity: Methods and
applications}. arXiv:201201976 {[}Cs{]} 2020.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Biecek2021}{}}%
\CSLLeftMargin{{[}39{]} }%
\CSLRightInline{Biecek P, Burzykowski T.
\href{https://ema.drwhy.ai/}{Explanatory {Model Analysis}: {Explore},
{Explain}, and {Examine Predictive Models}}. {New York}: {Chapman and
Hall/CRC}; 2021.}

\leavevmode\vadjust pre{\hypertarget{ref-Molnar2023}{}}%
\CSLLeftMargin{{[}40{]} }%
\CSLRightInline{Molnar C.
\href{https://christophm.github.io/interpretable-ml-book}{Interpretable
{Machine Learning}: {A Guide} for {Making Black Box Models
Explainable}}. 2023.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Yang2006}{}}%
\CSLLeftMargin{{[}43{]} }%
\CSLRightInline{Yang L.
\href{http://www.cs.cmu.edu/~./liuy/frame_survey_v2.pdf}{Distance
{Metric Learning}: {A Comprehensive Survey}}. PhD thesis. Michigan State
University, 2006.}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Xing2002}{}}%
\CSLLeftMargin{{[}45{]} }%
\CSLRightInline{Xing E, Jordan M, Russell SJ, Ng A.
\href{https://proceedings.neurips.cc/paper/2002/hash/c3e4035af2a1cde9f21e1ae1951ac80b-Abstract.html}{Distance
{Metric Learning} with {Application} to {Clustering} with
{Side-Information}}. Advances in {Neural Information Processing
Systems}, vol. 15, {MIT Press}; 2002.}
